\documentclass[12pt, fullpage,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{graphicx}

\newcommand{\semester}{Spring 2018}
\newcommand{\assignmentId}{0}
\newcommand{\releaseDate}{10 January, 2018}
\newcommand{\dueDate}{19 January, 2018}

\newcommand{\bx}{{\bf x}}
\newcommand{\bw}{{\bf w}}

\title{CS 5350/6350: Machine Learining \semester}
\author{Homework \assignmentId}
\date{Handed out: \releaseDate\\
  Due date: \dueDate}

\begin{document}
\maketitle

\input{emacscomm}
\input{general-instructions}


\section*{Basic Knowledge}
\label{sec:q1}

%problem 1, decide wether dependency & indendency
%2, prove p(A + B) <= P(A) + P(B)
%3, prove P(\sum_i A_i) \le \sum_i p(A_i)
%4. given two a joint Gaussian, calculate the conditional Guassian distribution
%5. prove E(X) = E(E(X|Y))
%4, prove V(E(X)) = E(V(X))
%5. prove V(Y) = EV(Y|X) + VE(Y|X)

%independency, conditional distribution, expectation, variance, basic properties
%gradient calcualtion, logistic function, second derivatives
%
\begin{enumerate}
\item~[5 points] We use sets to represent events. For example, toss a fair coin $10$ times, and the event can be represented by the set of ``Heads" or ``Tails" after each tossing. Let a specific event $A$ be ``at least one head". Calculate the probability that event $A$ happens, i.e., $p(A)$.
\item~[5 points] Given two events $A$ and $B$, prove that 
\[
p(A \cup B) \le p(A) + p(B).
\]
When does the equality hold?
\item~[10 points] Let $\{A_1, \ldots, A_n\}$ be a collection of events. Show that
\[
p(\cup_{i=1}^n A_i) \le \sum_{i=1}^n p(A_i).
\]
When does the equality hold?
\item~[5 points] Given three events $A$, $B$ and $C$, show that
\[
p(A\cap B\cap C) = p(A|B\cap C)p(B|C)p(C)
\]
\item~[20 points]  We use $\EE(\cdot)$ and $\VV(\cdot)$ to denote a random variable's mean (or expectation) and variance, respectively. Given two discrete random variables $X$ and $Y$, where $X \in \{0, 1\}$ and $Y \in \{0,1\}$. The joint probability $p(X,Y)$ is given in as follows:
\begin{table}[h]
        \centering
        \begin{tabular}{ccc}
        \hline\hline
         & $Y=0$ & $Y=1$ \\ \hline
         $X=0$ & $1/10$ & $2/10$ \\ \hline
         $X=1$  & $3/10$ & $4/10$ \\ \hline\hline
        \end{tabular}
        %\caption{Training data for the alien invasion problem.}\label{tb-alien-train}
        \end{table}
	
        \begin{enumerate}
            \item~[10 points] Calculate the following distributions and statistics. 
            \begin{enumerate}
            \item the the marginal distributions $p(X)$ and $p(Y)$
            \item the conditional distributions $p(X|Y)$ and $p(Y|X)$
            \item $\EE(X)$, $\EE(Y)$, $\VV(X)$, $\VV(Y)$
            \item  $\EE(Y|X=0)$, $\EE(Y|X=1)$,  $\VV(Y|X=0)$, $\VV(Y|X=1)$ 
            \item  the covariance between $X$ and $Y$
            \end{enumerate}
            \item~[5 points] Are $X$ and $Y$ independent? Why?
            \item~[5 points] When $X$ is not assigned a specific value, are $\EE(Y|X)$ and $\VV(Y|X)$ still constant? Why?
        \end{enumerate}
\item~[10 points] Assume a random variable $X$ follows a standard normal distribution, \ie $X \sim \N(X|0, 1)$. Let $Y = e^X$. Calculate the mean and variance of $Y$.
\begin{enumerate}
	\item $\EE(Y)$
	\item $\VV(Y)$
\end{enumerate}

\item~[10 points]  Given two random variables $X$ and $Y$, show that 
\begin{enumerate}
\item $\EE(\EE(Y|X)) = \EE(Y)$
\item
$\VV(Y) = \EE(\VV(Y|X)) + \VV(\EE(Y|X))$
\end{enumerate}

\item~[20 points]  Let us go back to the coin tossing example. Suppose we toss a coin for $n$ times, independently. Each toss we have a $0.3$ chance to obtain the head. Let us denote the total number of heads by $c(n)$. Calculate and derive the following statistics. 
\begin{enumerate}
\item $\EE(c(1))$, $\VV(c(1))$
\item $\EE(c(10))$, $\VV(c(10))$
\item $\EE(c(n))$, $\VV(c(n))$
\end{enumerate} 
What can you conclude from comparing the expectations and variances with different choices of $n$?  

\item~[10 points]   Given a logistic function, $f(\x) = 1/(1+\exp(-\a^\top \x))$ ($\x$ is a vector), derive/calculate the following gradients and Hessian matrices.  
\begin{enumerate}
\item $\nabla f(\x)$
\item $\nabla^2 f(\x)$
\item $\nabla f(\x)$ when $\a = [1,1,1,1,1]^\top$ and $\x = [0,0,0,0,0]^\top$
\item $\nabla^2 f(\x)$  when $\a = [1,1,1,1,1]^\top$ and $\x = [0,0,0,0,0]^\top$
\end{enumerate}

\item~[5 points] Show that for all $x \in \mathbb{R}$, 
\[1+x \le e^x, \;\;\;\;\;\; 1-x \le e^{-x}.
\]

\end{enumerate}


\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
